R Notebook

### 2020-03-28 datasets/data_preprocessing.nb.Rmd
update datasets - add virus dataset - add keywords for hyperparameter optimization - add datasets without keywords for simulation

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
This notebook processes all five systematic review datasets into datasets that can be used for testing asreview.

```{r, results = "hide"}
library(tidyverse)
library(knitr)
library(kableExtra)
```

### Set-up
```{r}
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

@@

# statistics on duplicates and dropping in ASReview dataset
drops <- data.frame(n = rep(0,5), # candidates
                    na = rep(0,5),# number of missing abstracts
                    narate = rep(0,5), # percenta of missing abstracts
                    dup = rep(0,5),  # number of duplicate abstracts 
                    duprate = rep(0,5), # percentage of missing abstracts
                    row.names = c("ace", "nudging", "ptsd", "software", "wilson"))

@@

## Ace dataset 
```{r}
ace <- template
# paper ------------------------------------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

@@

# asreview ---------------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)

@@

### Create test dataset
```{r}
# replace empty abstracts by NA values. 
ace_asr[ace_asr$abstract == "", "abstract"] <- NA

# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

# select columns
ace_test <- ace_test %>%
       select(authors, title, abstract, label_included) # select relevant columns

@@

### Some statistics 
```{r}
# missingness and duplicates in raw data on ASReview
drops["ace",] <- ace_asr %>%
          summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract))
drops["ace",]
ace
```

@@

```{r eval = FALSE}
# data is not public yet

# raw --------------------------------------------------------------------------
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")
# asreview ---------------------------------------------------------------------

# remove missing abstracts 
nudging_test <- n_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts
# missing statistics

@@

```{r}
nudging["test", ] <- data_descr(nudging_test)
```

@@

@@

### Create test dataset
Now, remove duplicate entries and empty abstracts to arrive at test set. 
```{r}
ptsd_test <- ptsd_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "") %>%
  distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts
```

@@

```{r}
# finalize dataset  
ptsd_test <- ptsd_test %>%
  select(authors, title, abstract, included, inclusion_code) # select relevant columns
```

Derive statistics in final test set.
```{r}
ptsd["test", ] <- data_descr(ptsd_test)

# statistics on duplicates from asr data to test
drops["ptsd", ] <- ptsd_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract))

ptsd
```

@@

### Statistics 
```{r}
drops["software", ] <- 
  hall_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract))

@@

@@

# add inclusion_code label 
w_asr$inclusion_code <- w_asr$label_abstract_screening + w_asr$label_included
w_asr$included <- w_asr$label_included
wilson["asreview", ] <- data_descr(w_asr)
```

```{r}
# create test dataset
w_test <- w_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "",
         abstract!= " ") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

# select relevant columns
w_test <- w_test %>%
  select(authors, title, abstract, included, inclusion_code)

@@

Statistics
```{r}
drops["wilson", ] <- w_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)), # todo correct this, "NA", or ""
                    narate = sum(is.na(abstract))/length(abstract),
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract))

@@

# save datafile, to serve as data for descriptive table in manuscript
save(all, file = "../manuscript/drafts/data/datasets_table.Rdata")
```

@@

Descriptives on missingness and duplicate abstracts in the ASReview test data set. 
```{r, echo = FALSE}
drops %>%
  kable("html",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2) %>%
  kable_styling(full_width = TRUE)
```

# Write all test dataset files 
```{r}
# ace
write.csv(ace_test, "test_datasets/ace.csv")

# nudging 
write.csv(nudging_test, "test_datasets/nudging.csv")

# ptsd
write.csv(ptsd_test, "test_datasets/ptsd.csv")

# hall
write.csv(hall_test, "test_datasets/software.csv")
# wilson
write.csv(w_test, "test_datasets/wilson.csv")

@@

>>>>

---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
This notebook processes all systematic review datasets into datasets suitable testing asreview.

Every dataset is a .csv file with the following columns: 
  - title
  - abstract
  - label_included
  - keywords? (?)

Entries with missing abstracts are removed

```{r, include = FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
```

### Set-up
```{r, echo = FALSE}
datasets <- c("ace", "nudging", "ptsd", "software", "virus", "wilson")
nd <- length(datasets)
# create storage for statistics
template <- data.frame(search = rep(0,4), # papers obtained in systematic search
                       ftext = rep(0, 4), # number of abstract inclusions
                       incl = rep(0,4), # papers included in final systematic review

@@

# statistics on duplicates and dropping in ASReview dataset
drops <- data.frame(n = rep(0,nd), # candidates
                    na = rep(0,nd),# number of missing abstracts
                    narate = rep(0,nd), # percentage of missing abstracts
                    dup = rep(0,nd),  # number of duplicate abstracts 
                    duprate = rep(0,nd), # percentage of missing abstracts
                    row.names = datasets)

@@

## ACE
```{r}
ace <- template
# paper ------------------------------------------------------------------------
ace["paper", ] <- c(2544, NA, round(2544*0.0160))

@@

# asreview ---------------------------------------------------------------------
ace_asr <- read.csv("https://raw.github.com/asreview/systematic-review-datasets/master/datasets/Cohen_EBM/output/ACEInhibitors.csv",
                     header=T)
# replace empty abstracts by NA values.

ace_asr[ace_asr$abstract == "", "abstract"] <- NA
# stats on missingness in this data
drops["ace",] <- ace_asr %>%
           summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

@@

### Create test dataset
```{r}
# drop missings and duplicates. 
ace_test <- ace_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

@@

### Some statistics 
```{r}
# missingness and duplicates in raw data on ASReview
drops["ace",]
ace
```

@@

```{r eval = TRUE}
# data is not public yet

# raw --------------------------------------------------------------------------
n_raw <- read.csv("../../scripting/nagtegaal_preprocessing/data/output/nagtegaal.csv")

# missing statistics 
drops["nudging",] <- n_raw %>%
          summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    na_rate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    dup_rate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
# asreview ---------------------------------------------------------------------

# remove missing abstracts 
nudging_test <- n_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

nudging["test", ] <- data_descr(nudging_test)

@@

### Some statistics
```{r}

drops["nudging",]
nudging
```

@@

# statistics on duplicates from asr data to test
drops["ptsd", ] <- ptsd_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
```

### Create test dataset
Now, remove duplicate entries and empty abstracts to arrive at test set. 
```{r}
ptsd_test <- ptsd_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  filter(abstract != "", abstract != "NA") %>%
  distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

@@

@@

```{r}
# finalize dataset  
ptsd_test <- ptsd_test %>%
  select(authors, title, abstract, keywords, included, inclusion_code) # select relevant columns

ptsd["test", ] <- data_descr(ptsd_test)
```

Derive statistics in final test set.
```{r}
drops["ptsd", ]
ptsd
```

@@

### Statistics 
```{r}
drops["software", ] <- 
  hall_asr %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

@@

## Virus dataset
```{r}
virus <- template
virus["paper",] <- c(2481, 132, 120)
v_raw <- read.csv("../../sims/data/virus.csv")
v_raw$abstract[v_raw$abstract == ""] <- NA

virus["raw", ] <- data_descr(v_raw)

drops["virus", ] <- 
  v_raw %>%
  summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)

v_test <- v_raw %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
   filter(abstract != "") %>%
   distinct(abstract, .keep_all = TRUE)# remove entries with duplicate abstracts

virus["test",] <- data_descr(v_test)
```

@@

w_asr[w_asr$abstract == "", "abstract"] <- NA
drops["wilson", ] <- w_asr %>%
   summarise(n = length(abstract),
                    na = sum(is.na(abstract)),
                    narate = sum(is.na(abstract))/length(abstract)*100,
                    dup = sum(duplicated(abstract, incomparables = NA)),
                    duprate = sum(duplicated(abstract, incomparables = NA))/length(abstract)*100)
# add inclusion_code label 
w_asr$inclusion_code <- w_asr$label_abstract_screening + w_asr$label_included
w_asr$included <- w_asr$label_included
wilson["asreview", ] <- data_descr(w_asr)
```

```{r}
# create test dataset
w_test <- w_asr %>%
  drop_na(abstract) %>% # drop entries with missing abstracts
  distinct(abstract, .keep_all = TRUE) # remove entries with duplicate abstracts

@@

Statistics
```{r}
drops["wilson",]

@@

# save datafile, to serve as data for descriptive table in manuscript
save(all, file = "../manuscript/drafts/data/datasets_table.Rdata")

saveRDS(drops, file = "data_statistics/drops.RDS")
```

@@

Descriptives on missingness and duplicate abstracts in the ASReview test data set. 
```{r, echo = FALSE}
drops %>%
  kable("html",
        col.names = c("n", "NA", "NA rate (%)", "duplicates", "duplicate rate (%)"),
        digits = 2) %>%
  kable_styling(full_width = TRUE)

@@

# Write all test dataset files 
```{r}
# ace
write.csv(ace_test %>%  select(authors, title, abstract, keywords, label_included), "test_datasets/ace.csv")

# nudging (no keywords available)
write.csv(nudging_test %>% select(title, abstract, included), "test_datasets/nudging.csv")

# ptsd
write.csv(ptsd_test %>% select(authors, title, abstract, keywords, included), "test_datasets/ptsd.csv")

# hall (no keywords available )
write.csv(hall_test, "test_datasets/software.csv")

# virus 
write.csv(v_test %>% select(authors, title, abstract, keywords, included), "test_datasets/virus.csv")

# wilson
write.csv(w_test %>% select(authors, title, abstract, keywords, included), "test_datasets/wilson.csv")
```

# write test dataset files for simulation (removing keywords)
```{r}
# ace
write.csv(ace_test %>%  select(authors, title, abstract, label_included), "sim_datasets/ace.csv")
# nudging (no keywords available)
write.csv(nudging_test %>% select(title, abstract, included), "sim_datasets/nudging.csv")
# ptsd
write.csv(ptsd_test %>% select(authors, title, abstract, included), "sim_datasets/ptsd.csv")
# hall (no keywords available )
write.csv(hall_test, "sim_datasets/software.csv")
# virus 
write.csv(v_test %>% select(authors, title, abstract, included), "sim_datasets/virus.csv")
# wilson
write.csv(w_test %>% select(authors, title, abstract, included), "sim_datasets/wilson.csv")
```
